---
title: "ibdrel-r"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ibdrel-r}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Basic usage

The default model, `ibdrel-unilineal` is loaded using the `fitModel()` function, with `data = NULL`. Features and match threshold values are set using the `features` and `cutoff` arguments. In the following example, the default model is used to classify an artifical example of IBD sharing (segment lengths).

```{r setup}
library(ibdrel)

model = fitModel()


```

### Example:


### Example: Simulated SNP data from a pair of first cousins

As a second example 

```{r comment=NA, warning=F}
library(pedtools)
library(ibdsim2)
library(ibdfindr)

x = cousinPed(1) |> distributeMarkers(n = 50000)
sim = ibdsim(x, seed = 123, verbose = FALSE)
y = profileSimIBD(x, sim, ids = leaves(x), seed = 123)

ibd = findIBD(y)
```

Segments can be converted to a list compatible with `ibdrel` using the function `ibdfindr2ibdrel()`.


```{r}
segments = ibdfindr2ibdrel(ibd)

pred = predict(segments, model)

pred

```



## Building
