---
title: "Aggregating Classes"
author: "Henrik Nordtorp"
date: "`r Sys.Date()`"
output: html_document
---

This notebook explores the effects of aggregating classification results. Should classes/data be aggregated prior to classification, or should posterior probabilities be aggregated after classification? 

```{r}
library(ibdrel)
library(pedtools)
library(ibdsim2)
library(dplyr)

# Training data at the "relationship" level

segmentDataRel = readRDS(system.file("data", "segments_unilineal.rds", package = "ibdrel"))

# Pedigrees with associated metadata
pedsDataRel = readRDS(system.file("data", "peds_unilineal.rds", package = "ibdrel"))
pedsMetadataRel = pedigreesMetadata(pedsDataRel)
```

Create a common "test" observation:

```{r}
set.seed(1234)
test <- runif(100, min = 1, max = 50)
test
```



Classify the data "as is":
```{r}
pdfuns_rel = lapply(segmentDataRel, preparePdfs)
pred_rel = classify(test, pdfuns_rel)
res.df.tmp = as.data.frame(pred_rel)
res.df.tmp$Relationship = rownames(res.df.tmp)

res_deg = merge(res.df.tmp, pedsMetadataRel, by = "Relationship",
                all.x = T, all.y = F)
```

Classify the data at the "Donnelly" level:
```{r}
segmentDataClass = list()

for (i in 1:length(segmentDataRel)) {
  rel = names(segmentDataRel)[i]
  don = pedsMetadataRel |> 
    filter(Relationship == rel) |> 
    select(class) |> as.character()
  
  data.rel = segmentDataRel[[i]]
  
  for (segment in data.rel) {
    if (don %in% names(segmentDataClass)) {
      segmentDataClass[[don]][[length(segmentDataClass[[don]])+1]] <- segment
    } else {
      segmentDataClass[[don]][[1]] <- segment
    }
  }
    
}

pdfuns_don = lapply(segmentDataClass, preparePdfs)
pred_don = classify(test, pdfuns_don)
```

Classify data at the "degree" level

```{r}
segmentDataDeg <- list()

for (i in 1:length(segmentDataRel)) {
  rel = names(segmentDataRel)[i]
  deg = pedsMetadataRel |> 
    filter(Relationship == rel) |> 
    select(degree) |> as.character()
  
  data.rel = segmentDataRel[[i]]
  
  for (segment in data.rel) {
    if (deg %in% names(segmentDataDeg)) {
      segmentDataDeg[[deg]][[length(segmentDataDeg[[deg]])+1]] <- segment
    } else {
      segmentDataDeg[[deg]][[1]] <- segment
    }
  }
    
}

pdfuns_deg = lapply(segmentDataDeg, preparePdfs)
pred_deg = classify(test, pdfuns_deg)
```


Results

At the degree level

```{r}
res_deg |> 
  group_by(degree) |> 
  summarise(p = sum(pred_rel)) -> res_deg
pred_deg <- as.data.frame(pred_deg)
pred_deg$degree <- rownames(pred_deg)

df = merge(res_deg, pred_deg,
           by.x = "degree", by.y = "degree")
df = df |> 
  pivot_longer(-degree, names_to="source", values_to="p")

library(tidyverse)
ggplot(df)+
  geom_col(aes(x = degree, y = p, fill = source), position = "dodge")
```

